---
title: "Summary visualisation of cingulotomy cell densities"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(forcats)
library(ggdark)

options(scipen=999)
theme_set(dark_theme_bw(base_size = 14) + theme(legend.position = "top",
                                                panel.grid = element_blank()))
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 9, 
                      fig.height = 4)

cust_palette <- c("white", "#56B4E9")
```

```{r load-data}
path_glia <- here("Data", "Cingulotomy_brains_analysis_NeuN-GFAP-Iba1.rds")
df_glia <- read_rds(path_glia) %>%
  dplyr::mutate(Region = ROI)

path_GABA <- here("Data", "Cingulotomy_brains_analysis_NeuN-GABA-GAD67.rds")
df_GABA <- read_rds(path_GABA) %>%
  dplyr::mutate(Region = ROI)

path_wm <- here("Data", "White-matter-measurements.rds")
df_wm <- read_rds(path_wm) %>% dplyr::mutate(Treatment = 
                                               fct_recode(Treatment, 
                                                      ACC = "cingulotomy",
                                                      CON = "control"), 
                                       Treatment = 
                                       fct_relevel(Treatment, "CON"))
```

```{r aggregate-ROIs-glia}
df_glia_r <- df_glia %>% dplyr::select(Animal_ID, Treatment, Region, Antibody, 
                                       Slide, Section, Num_detections, 
                                       Area_sq_mm) %>% 
  dplyr::arrange(Animal_ID, Slide, Section, Region, Antibody) %>% 
  dplyr::mutate(NeuN_detections = ifelse(Antibody == "NeuN", 
                                         Num_detections, NA_real_)) %>% 
  tidyr::fill(NeuN_detections) %>% 
  dplyr::mutate(Region = dplyr::recode_factor(Region, 
                                  Area_24 = "ACC", 
                                  Area_23 = "ACC", DLPAG = "PAG",
                                  DMPAG = "PAG", LPAG = "PAG", VLPAG = "PAG",
                                  Amy_AB = "BAm", Amy_B = "BAm", 
                                  .default = levels(Region)),
                Treatment = fct_recode(Treatment, 
                                                 ACC = "cingulotomy",
                                                 CON = "control")) %>% 
  dplyr::group_by(Animal_ID, Treatment, Region, Antibody) %>% 
  dplyr::summarise(Cells_per_sq_mm = sum(Num_detections) / sum(Area_sq_mm),
                   NeuN_detections = first(ifelse(Antibody == "NeuN", 
                                            sum(NeuN_detections) / sum(Area_sq_mm),
                                            NA_real_))) %>% 
  dplyr::arrange(Animal_ID, Region, Antibody) %>% 
  tidyr::fill(NeuN_detections) %>% ungroup() %>% 
  dplyr::mutate(Prop_nonNeuN_NeuN = Cells_per_sq_mm / (Cells_per_sq_mm + NeuN_detections))

# path_glia_r <- here("Data", "Cingulotomy_brains_analysis_NeuN-GFAP-Iba1_summary.rds")
# write_rds(df_glia_r, path_glia_r)
```
```{r aggregate-ROIs-gaba}
df_GABA_r <- df_GABA %>% dplyr::select(Animal_ID, Treatment, Region, Antibody, 
                                       Slide, Section, Num_detections, 
                                       Area_sq_mm) %>% 
  dplyr::arrange(Animal_ID, Slide, Section, Region, Antibody) %>% 
  dplyr::mutate(NeuN_detections = ifelse(Antibody == "NeuN", 
                                         Num_detections, NA_real_)) %>% 
  tidyr::fill(NeuN_detections) %>% 
  dplyr::mutate(Region = dplyr::recode_factor(Region, 
                                  Area_24 = "ACC", 
                                  Area_23 = "ACC", DLPAG = "PAG",
                                  DMPAG = "PAG", LPAG = "PAG", VLPAG = "PAG",
                                  Amy_AB = "BAm", Amy_B = "BAm", 
                                  .default = levels(Region)),
                Treatment = fct_recode(Treatment, 
                                                 ACC = "cingulotomy",
                                                 CON = "control")) %>% 
  dplyr::group_by(Animal_ID, Treatment, Region, Antibody) %>% 
  dplyr::summarise(Cells_per_sq_mm = sum(Num_detections) / sum(Area_sq_mm),
                   NeuN_detections = first(ifelse(Antibody == "NeuN", 
                                            sum(NeuN_detections) / sum(Area_sq_mm),
                                            NA_real_))) %>% 
  dplyr::arrange(Animal_ID, Region, Antibody) %>% 
  tidyr::fill(NeuN_detections) %>% ungroup() %>% 
  dplyr::mutate(Prop_nonNeuN_NeuN = Cells_per_sq_mm / (Cells_per_sq_mm + NeuN_detections))

# path_GABA_r <- here("Data", "Cingulotomy_brains_analysis_NeuN-GABA-GAD67_summary.rds")
# write_rds(df_GABA_r, path_GABA_r)
```
**Figure 1**. Proportion of astrocytes to the sum of astrocytes and NeuN-positive cells in five brain regions conditioned on Treatment. 
```{r astro-graph}
g_astro <- ggplot(df_glia_r %>% dplyr::filter(Antibody == "GFAP"),
                    aes(x = Treatment,
                        y = Prop_nonNeuN_NeuN,  colour = Treatment)) + 
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. - 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. + 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  geom_point(size = 3, position = position_jitter(width = 0.2)) + 
  scale_colour_manual(values = cust_palette) +
  facet_wrap(~Region, nrow = 1) +
  labs(y = "Proportion of astrocytes", x = NULL) 
g_astro
```

**Figure 2**. Proportion of microglia to the sum of microglia and NeuN-positive cells in five brain regions conditioned on Treatment. 
```{r microg-graph}
g_microg <- ggplot(df_glia_r %>% dplyr::filter(Antibody == "Iba1"),
                    aes(x = Treatment,
                        y = Prop_nonNeuN_NeuN,  colour = Treatment)) + 
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. - 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. + 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  geom_point(size = 3, position = position_jitter(width = 0.2)) +
  scale_colour_manual(values = cust_palette) +
  facet_wrap(~Region, nrow = 1) +
  labs(y = "Proportion of microglia", x = NULL) 
g_microg
```

**Figure 3**. Proportion of GABA-positive cells to the sum of GABA-positive cells and NeuN-positive cells in five brain regions conditioned on Treatment. 
```{r gaba-graph}
g_GABA <- ggplot(df_GABA_r %>% dplyr::filter(Antibody == "GABA"),
                    aes(x = Treatment,
                        y = Prop_nonNeuN_NeuN,  colour = Treatment)) + 
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. - 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. + 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  geom_point(size = 3, position = position_jitter(width = 0.2)) + 
  scale_colour_manual(values = cust_palette) +
  facet_wrap(~Region, nrow = 1) +
  labs(y = "Proportion of GABA-positive cells", x = NULL)
g_GABA
```

**Figure 4**. Cross-sectional diameters of white matter fibre bundles in six brain regions conditioned on Treatment. 
```{r wm-graph, fig.height = 6}
g_wm <- ggplot(df_wm,
                    aes(x = Treatment,
                        y = Length_um,  colour = Treatment)) + 
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. - 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  stat_summary(fun = "mean", geom = "segment", mapping = aes(xend = ..x.. + 0.3, 
                                                             yend = ..y..),
               color = "gray60", size = 1.6) +
  geom_point(size = 3, position = position_jitter(width = 0.2)) + 
  scale_colour_manual(values = cust_palette) +
  facet_wrap(~ROI, nrow = 2) +
  labs(y = "Diameter of white matter bundles", x = NULL)
g_wm
```